---
title: "Correlation between covariates"
author: "Edith"
date: "`r format(Sys.time(), '%d %m %Y')`"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: true
    toc: true
---

```{r, message=FALSE, warning=FALSE}
# Load necessary libraries
library(kableExtra)
source("../../../R_helpers/generic.R")
cov_df <- read_csv(file.path(out_dir, "covariates", "final",  "ua_covariates_oblast.csv")) %>%
  mutate(
         covariate = str_replace(covariate, 'disorder', "dis"),
         covariate = str_replace(covariate, 'Demonstration', "Dem"),
    scale_label = paste0(sum_stat, ", ", time_std, ", ", space_std),
    cov_process = paste0(covariate, ",", scale_label)
    )
master_index <- read_csv(file.path(out_dir, paste0(tolower(country), "_master_index.csv")))|>
  distinct(i_key, i_name, i, ADM1_PCODE, macroregion)
cov_df <- cov_df |> 
  left_join(master_index)
```

This report illustrates the correlation between covariates. Because there are `r n_distinct(cov_df |> pull(cov_process))` covariates, it splits the correlation analysis into three parts:

1.  Correlation between processing methods for each covariate

2.  Correlation between covariates for each processing method

3.  Full correlation table for further exploration.

The correlation is calculated using the Pearson method between covariate values computed for each oblast and time point.

# Correlation between processing methods

```{r}
#| output: asis

for(cov in sort(unique(cov_df$covariate), decreasing = T)){
  
  cat(paste0("\n\n## ", cov, "\n\n"))
  
  cov_corr_plot <- cov_df |> 
  filter(covariate == cov) |> 
  select(i, t, cov_process, value_std) |> 
  pivot_wider(names_from = cov_process, values_from = value_std) |> 
  select(-i, -t) %>%
  cor(use = "pairwise.complete.obs") |> 
  as.matrix() |> 
    corrplot::corrplot(method = "color", type = "upper", tl.col = "black", tl.srt = 45)
  
  cat("\n")
}
```

# Correlation between covariates per processing method

```{r}
#| output: asis
#| 
for(process in unique(cov_df$scale_label)){
  
  cat(paste0("\n\n## ", process, "\n\n"))
  process_corr_plot <- cov_df |> 
  filter(scale_label == process) |> 
  mutate(covariate = str_replace(covariate, 'acled_', "A")) |> 
  select(i, t, covariate, value_std) |> 
  pivot_wider(names_from = covariate, values_from = value_std) |> 
  select(-i, -t) %>%
  cor(use = "pairwise.complete.obs") |> 
  as.matrix() |> 
    corrplot::corrplot(method = "color", type = "upper", tl.col = "black", tl.srt = 45)
  
  cat("\n")
}
```

# Full correlation table

```{r}
cov_df_corr <- cov_df %>%
  select(i, t, cov_process, value_std) |> 
  pivot_wider(names_from = cov_process, values_from = value_std) |> 
  select(-i, -t) %>%
  cor(use = "pairwise.complete.obs") |> 
  as.data.frame() |> 
  rownames_to_column(var = "cov_process") |> 
  pivot_longer(-cov_process, names_to = "cov_process2", values_to = "correlation") |>
  filter(correlation <0.99) |> 
  rowwise() |> 
  mutate(correlation = round(correlation, 2),
         type = factor(case_when(
           str_split(cov_process, ",")[[1]][1] == str_split(cov_process2, ",")[[1]][1] ~ "same covariate",
           grepl('acled', str_split(cov_process, ",")[[1]][1]) & grepl('acled', str_split(cov_process2, ",")[[1]][1]) ~ "same covariate",
           str_split_fixed(cov_process, ",", 2)[[2]] == str_split_fixed(cov_process2, ",", 2)[[2]] ~ "same processing",
           T ~ "different processing and covariate"
         ))
         ) 

cov_df_corr |> 
  select(type, cov_process, cov_process2, correlation) |>
  DT::datatable(rownames = FALSE, options = list(pageLength = 10, scrollX = TRUE), filter="top") 
```
